---
title: "L1morphPELIC"
author: "Daniel Crawford"
date: "04/02/2024"
output: 
  github_document: 
    toc: TRUE
---
EXISTING

# Correlation of Future Tense Construction Preference with Proficiency Scores for English L2 Learners

```{r}
#Import Packages
library(tidyverse)
```





### Load Data
Raw data can be found [here](https://github.com/ELI-Data-Mining-Group/PELIC-dataset/raw/master/PELIC_compiled.csv).

```{r}
#Read in Data from PELIC: 
#PELIC_compiled = as_tibble(read.csv(url("https://github.com/ELI-Data-Mining-Group/PELIC-dataset/raw/master/PELIC_compiled.csv"), fileEncoding = "ISO-8859-1"))
PELIC_compiled = as_tibble(read.csv('data/pelic_compiled.csv'))

#Need to get demographic and proficiency scores for each student

student_info = as_tibble(read.csv(url("https://github.com/ELI-Data-Mining-Group/PELIC-dataset/raw/master/corpus_files/student_information.csv")))


#Read in tokenized df:
tokenized_df = readRDS('data/tokenized_df.rds')


#If you dont have the tokenized df, this will create it in about 10 minutes:

#A function to create a tokenized df with columns: token | lemma | POS
#Input is a string
tlp_text_to_df = function(x){
  text_df = x %>% 
    #Remove "[" character (at beginning of string)
    str_remove("^\\[\\(") %>% 
    #Remove "]" character (at end of string)
    str_remove("\\)\\]$") %>% 
    #split tuples
    str_split_1("\\), \\(") %>% 
    #remove 's
    str_remove_all("'") %>% 
    #convert to tibble
    as.tibble() %>% 
    #remove rows of commas - these are not relevant or impactful on analysis
    filter(!startsWith(value, ",,")) %>% 
    #separate into columns
    separate(value, into = c('token','lemma','POS'), sep = ',')
  
  return(text_df)
  
}


#tokenized_df =  PELIC_compiled %>% mutate(tokenized_nested_df = map(tok_lem_POS, tlp_text_to_df))
  
  
tokenized_df

```


# Set Lemmas



```{r}
#Set lemmas: {lemma}_{N/V}
lemmas = c('walk_V','help_V','work_V')




```

```{r}

#Tokenized df with just lemmas matching
tokenized_df_just_lemmas = tokenized_df %>%
  select(answer_id, tokenized_nested_df) %>% 
  unnest(tokenized_nested_df) %>% 
  mutate(lemma_and_pos = paste(trimws(lemma), substr(trimws(POS),1,1), sep = '_')) %>% 
  filter(lemma_and_pos %in% lemmas) %>% 
  select(-lemma_and_pos)

tokenized_df_just_lemmas
```




### Ben's Demo Data
Columns for 
-   count of base forms of a word by student -> count_base_forms
-   count of -s inflections, -ing, -ed -> ending with _
-   count of total inflected words -> count_total_inflected_words
-   number of different inflections (up to 4) -> count_number_differnt_inflections
-   ratio of inflected to lemmas







```{r}

#Create tokenized_df, each row is a token
count_base_forms = tokenized_df_just_lemmas %>% 
  filter(trimws(token)==trimws(lemma)) %>% 
  group_by(answer_id, lemma) %>% 
  summarise(base_form_count = n()) %>% 
  ungroup(answer_id, lemma) %>% 
  pivot_wider(names_from = lemma, names_prefix = "base_form_count", values_from = base_form_count, values_fill = 0)
  


count_base_forms

```


```{r}
count_ending_in = function(text_df, lemmas, str){
  
  text_df %>% 
    filter(endsWith(trimws(token), str)) %>% 
    group_by(answer_id, lemma) %>% 
    summarise(count_ending = n()) %>% 
    ungroup(answer_id, lemma) %>% 
    pivot_wider(names_from = lemma, names_prefix = paste0("endsWith_",str), values_from = count_ending, values_fill = 0)
  
}

ending_with_s = count_ending_in(tokenized_df_just_lemmas, lemmas, str = 's')
ending_with_ing = count_ending_in(tokenized_df_just_lemmas, lemmas, str = 'ing')
ending_with_ed = count_ending_in(tokenized_df_just_lemmas, lemmas, str = 'ed')

ending_with_s
```


```{r}
count_total_inflected_words = tokenized_df_just_lemmas %>% 
  group_by(answer_id, lemma) %>% 
  summarise(total_inflected_words = n()) %>% 
  ungroup(answer_id, lemma) %>% 
  pivot_wider(names_from = lemma, names_prefix = "total_inflected_words", values_from = total_inflected_words, values_fill = 0)
  


count_total_inflected_words
```




```{r}
inflection_rato = tokenized_df_just_lemmas %>% 
  mutate(is_inflected = as.logical(trimws(token) != trimws(lemma))) %>% 
  group_by(answer_id, lemma) %>% 
  summarise(infl_ratio = sum(is_inflected)/n()) %>% 
  ungroup(answer_id, lemma) %>% 
  pivot_wider(names_from = lemma, names_prefix = "inflection_ratio", values_from = infl_ratio, values_fill = 0)
  


inflection_rato
```


```{r}
final_data = PELIC_compiled %>% 
  select(answer_id, anon_id) %>% 
  left_join(count_base_forms, by = 'answer_id') %>% 
  left_join(ending_with_s, by = 'answer_id') %>% 
  left_join(ending_with_ing, by = 'answer_id') %>% 
  left_join(ending_with_ed, by = 'answer_id') %>% 
  left_join(count_total_inflected_words, by = 'answer_id') %>% 
  #left_join(count_number_differnt_inflections, by = 'answer_id') %>% 
  #left_join(inflection_rato, by = 'answer_id') %>% 
  select(-answer_id) %>% 
  arrange(anon_id) %>% 
  group_by(anon_id) %>% 
  summarise(across(everything(), ~sum(., na.rm = T))) %>% 
  left_join(
    PELIC_compiled %>% 
      select(answer_id, anon_id) %>% 
      left_join(inflection_rato, by = 'answer_id') %>% 
      select(-answer_id) %>% 
      arrange(anon_id) %>% 
      group_by(anon_id) %>% 
      summarise(across(everything(), ~mean(., na.rm = T))),
    by = 'anon_id' 
  ) %>% 
  left_join(student_info, by = 'anon_id')


final_data
```


```{r}
final_data %>% write.csv('final_data.csv')
```









left_join(count_base_forms, by = 'answer_id') %>% ### Format the Data 
```{r}



#Create tokenized_df, each row is a token
lemma_counts = PELIC_compiled %>% 
  
  #Optional Slice, included for time
  slice(1:1000) %>% 
  
  #create tokenized df from the tok_lem_POS string
  mutate(tokenized_nested_df = map(tok_lem_POS, tlp_text_to_df)) %>% 
  #keep the unique identifier of answer id
  select(answer_id, tokenized_nested_df) %>% 
  #make the df longer and unndest
  unnest(tokenized_nested_df) %>% 
  
  
  
  filter(trimws(lemma) %in% lemmas) %>% 

  
  distinct() %>% 
  
  group_by(answer_id, lemma) %>% 
  
  summarise(different_tokens = n()) %>% 
  
  left_join(
    PELIC_compiled %>% select(answer_id, anon_id)
    ) %>% 
  pivot_wider(names_from = lemma, names_prefix = "lemma_count", values_from = different_tokens, values_fill = 0) %>% 
  
  left_join(student_info, "anon_id")



lemma_counts

```



# demo vis
```{r}
lemma_counts %>% 
  ungroup() %>% 
  select(starts_with('lemma_count'), native_language) %>% 
  group_by(native_language) %>% 
  summarise_all(.funs = mean) %>% 
  pivot_longer(cols = starts_with('lemma_count')) %>% 
  separate(name, c(NA, 'lemma'), sep = ' ') %>% 
  
  filter(native_language %in% c('Korean','Turkish','Chinese')) %>% 
  

  ggplot(aes(native_language, value, fill = lemma))+
    geom_bar(stat = 'identity', position = 'dodge')+
    facet_wrap(~lemma)+
    ggtitle("Average Counts of Different Lemmas by L1") +
    xlab('Native Language (L1)') + ylab('Average Nu,ber of Different Lemmas') +
    theme_minimal()
```



# Session Info
```{r}
sessionInfo()
```